\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-08-05 popl-2019-journal-v/discussion.tex
[POPLjv] Moderate wording about revoked regions
[POPLjv] Rewrite stack allocation discussion

  We conjecture that \stktokens{} can support a limited form of continuations and exceptions.
  Specifically, the continuations and exceptions would be passed to the callee as sealed code capabilities.
  In order for the continuations and exceptions to work with the stack, they would have to be sealed with the return seals.
  \DDin{I'm not convinced that these limitations cannot be avoided. Perhaps we should keep this discussion more tentative?}
  This requirement puts two strong limitations on continuations and exceptions.
  First of all, they can be used by the immediate callee, but they cannot be used in subsequent calls made by the callee (assuming they use \stktokens{}).
  The problem is that the callee would have to take up part of the stack for their local stack frame.
  This would break up the return token that should be used with the continuation (or exception).
  If the callee really needs to pass on the continuation, then they would have to construct a wrapper that unravels the stack and restores the necessary return token.
  Second, a call that passes continuations or exceptions in this way has to pass the same continuations and exceptions in every call.
  The problem is that just like the return code capability, the continuation and exception capabilities are non-linear.
  This means that a callee can store them for subsequent calls preventing us from using different continuations and exceptions for different calls.

@@

\begin{jversion}
Programs that use a C-like calling convention often allow programs to pass a stack reference in a call either to provide a reference for a piece of memory to work on or to indicate where the result of the call should be placed.
\stktokens{} does not support this as it expects to get a stack pointer for the entire stack it handed out.
\dominique{Err... I don't see the problem.
  Passing stack references to callees should work fine, as long as the callee treats them linearly (i.e.\ doesn't try to copy them).
  }
However, if we have allocated part of the stack for a different purpose and handed out a reference for it, then the stack pointer cannot reference the same memory as the stack pointer is a linear capability.
\dominique{I don't understand this sentence.}
However, while it is not possible to allocate a reference for the stack, one can easily imagine that the stack could still be used to pass and return values.
\dominique{Or this one...}
\end{jversion}

>>>>

  We conjecture that \stktokens{} can support a limited form of continuations and exceptions.
  Specifically, the continuations and exceptions would be passed to the callee as sealed code capabilities.
  In order for the continuations and exceptions to work with the stack, they would have to be sealed with the return seals.
  \dominique{I'm not convinced that these limitations cannot be avoided. Perhaps we should keep this discussion more tentative?}
  This requirement puts two strong limitations on continuations and exceptions.
  First of all, they can be used by the immediate callee, but they cannot be used in subsequent calls made by the callee (assuming they use \stktokens{}).
  The problem is that the callee would have to take up part of the stack for their local stack frame.
  This would break up the return token that should be used with the continuation (or exception).
  If the callee really needs to pass on the continuation, then they would have to construct a wrapper that unravels the stack and restores the necessary return token.
  Second, a call that passes continuations or exceptions in this way has to pass the same continuations and exceptions in every call.
  The problem is that just like the return code capability, the continuation and exception capabilities are non-linear.
  This means that a callee can store them for subsequent calls preventing us from using different continuations and exceptions for different calls.

@@

\begin{jversion}
  Programming languages with a C-like calling convention often allow programs to pass stack references in calls.
  \stkotkens{} supports stack references but with a couple of caveats.
  First of all, the stack capability is linear, so all references to the stack have to be linear.
  This means that the callee has to treat references linearly.
  Next like the stack capability, the stack references must be given back to the caller on return, so they can reconstruct their original stack capability (which allows them to return). 
  Finally, the encapsulated local stack frame should be a continuous piece of memory which means that the stack allocations need to be next to the stack frame.
  An escape analysis could be used to statically determine where to put allocations and, in principle, the allocations could be reordered dynamically before a call.
  It is not clear to us to what extend existing C-code is covered by the stack references supported by \stktokens{}.
\end{jversion}
